---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# spainpolls

`spainpolls` is an R client for the Spain Electoral Polls API. It returns tidy
`tibble`s for elections, pollsters, parties, polls, long-format poll results,
and voting-intention time series.

The package queries the public read-only API documented in
`docs/r-package-handoff/`. It does not access PostgreSQL directly, scrape data,
or modify records.

## Installation

```{r}
# install.packages("pak")
pak::pak("")
```

## Configure the API URL

By default, `spainpolls` uses:

```{r}
spainpolls_api_url()
```

To use a local API:

```{r}
set_spainpolls_api_url("http://localhost:8000")
```

You can also set the global option directly:

```{r}
options(spainpolls.api_url = "http://localhost:8000")
```

## Covered Endpoints

The package covers these public endpoints:

- `GET /health`
- `GET /api/v1/elections`
- `GET /api/v1/pollsters`
- `GET /api/v1/parties`
- `GET /api/v1/polls`
- `GET /api/v1/polls/{poll_id}`
- `GET /api/v1/results`
- `GET /api/v1/timeseries`

`GET /api/v1/polls/{poll_id}/results` is not implemented by the API. Use
`get_poll_results()` to fetch `/api/v1/polls/{poll_id}` and extract the nested
results.

## Basic Use

```{r}
library(spainpolls)

elections <- get_elections(limit = 10)
pollsters <- get_pollsters(election_key = "congreso_espana_2027")
parties <- get_parties(election_key = "congreso_espana_2027")
```

## Polls

```{r}
polls <- get_polls(
  election_key = "congreso_espana_2027",
  limit = 10
)

poll <- get_poll(101)
poll_results <- get_poll_results(101)
```

## Long Results

```{r}
results <- get_results(
  election_key = "congreso_espana_2027",
  party = "psoe",
  collect_all = TRUE
)
```

## Time Series

```{r}
ts <- get_timeseries(
  election_key = "congreso_espana_2027",
  party = "psoe"
)
```

## Plotting

```{r}
library(ggplot2)

ggplot(ts, aes(date, vote_share, color = pollster_name)) +
  geom_point() +
  geom_line(aes(group = pollster_name), alpha = 0.4) +
  labs(
    title = "Voting intention for PSOE",
    x = NULL,
    y = "Vote share (%)"
  )
```

## Pagination

Paginated functions accept `limit`, `offset`, and `collect_all`.

```{r}
all_results <- get_results(
  election_key = "congreso_espana_2027",
  collect_all = TRUE
)
```

When `collect_all = TRUE`, the client starts at offset 0 and requests pages
until the API reports that all matching rows have been downloaded.

## Errors

The package validates common bad inputs before making HTTP requests, including
invalid `limit`, `offset`, dates, logical flags, `poll_id`, and `date_field`.
HTTP errors from the API are rethrown as clear R errors with `cli`.

## Spain Electoral Project

This package is part of the Spain Electoral Project ecosystem. Cite or link the
Spain Electoral Project when using the API data in public work.

## Current Limitations

- The API is read-only and unauthenticated in v1.
- Administrative endpoints are not included.
- `get_poll_results()` extracts nested results because the API does not expose a
  separate poll-results route.
- Sorting order is determined by the API and is not specified as a stable client
  contract.
