---
title: "Getting started with spainpolls"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with spainpolls}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r}
library(spainpolls)
```

## Elections, Pollsters and Parties

```{r}
elections <- get_elections(limit = 10)

pollsters <- get_pollsters(
  election_key = "congreso_espana_2027"
)

parties <- get_parties(
  election_key = "congreso_espana_2027"
)
```

## Polls

```{r}
polls <- get_polls(
  election_key = "congreso_espana_2027",
  limit = 20
)

one_poll <- get_poll(101)
one_poll_results <- get_poll_results(101)
```

## Results

`get_results()` returns one row per poll and party.

```{r}
results <- get_results(
  election_key = "congreso_espana_2027",
  collect_all = TRUE
)
```

## Time Series

```{r}
ts <- get_timeseries(
  election_key = "congreso_espana_2027",
  party = "psoe",
  date_field = "auto"
)
```

## Analysis with dplyr

```{r}
library(dplyr)

get_results(election_key = "congreso_espana_2027", collect_all = TRUE) |>
  filter(party_key %in% c("psoe", "pp", "vox", "sumar")) |>
  group_by(party_key) |>
  summarise(mean_vote = mean(vote_share, na.rm = TRUE), .groups = "drop")
```

## Plotting with ggplot2

```{r}
library(dplyr)
library(ggplot2)

ts <- get_timeseries(
  election_key = "congreso_espana_2027",
  date_field = "fieldwork_end", 
  party_key = c("psoe", "pp", "vox", "sumar"),
  collect_all = T, 
  include_manual_review = T
)

ts |> 
  ggplot(aes(x = date, y = vote_share, fill = color_hex, color = color_hex, group = short_name)) +
  geom_hline(yintercept = 0, alpha = .3) +
  geom_smooth(span = .1, alpha = .2) +
  geom_point(
    aes(x = date, y = vote_share), 
    data = plot_data,
    alpha = .1
  ) + 
  scale_color_identity(aesthetics = c("color", "fill")) +
  theme_minimal() +
  labs(x = NULL, y = "Vote share (%)")
```

## Pagination

All list endpoints are paginated. Use `limit` and `offset` for one page, or `collect_all = TRUE` to download all matching pages:

```{r}
all_polls <- get_polls(
  election_key = "congreso_espana_2027",
  collect_all = TRUE
)
```

When collecting all rows, `spainpolls` starts at offset 0 and stops when the API reports that all matching rows have been downloaded.
